Project Proposal

### 2024-11-17 NFL_IDS702Projectproposal.qmd
change the variable of nfl Q1

Outcome Variable: Pass Result (passResult) - This is a binary variable,
with categories such as "Complete" (C) and "Incomplete" (I), indicating
whether the pass was successfully completed.

Independent Variables:

**dropbackType**: The type of drop back after the snap by the QB
(Traditional, Designed Rollout, Scramble, Scramble Rollout, Designed
Rollout Left, Designed Rollout Right, Scramble Rollout Left, Scramble
Rollout Right, Designed Run, QB Draw, Rollout, text) (categorical)\
**dropbackDistance**: The distance the QB dropped back (yards) behind
the center after the snap (numeric)\
**timeToSack**: Time before the QB is sacked; this indicates the level
of pass protection (numeric) **rushLocationType**: Indicates where
defensive pressure might be coming from; outside pressure can force
quicker throws. (categorical)

Possible other variables:\
**pff_passCoverage**: Defensive coverage scheme (e.g., man vs. zone);
some schemes are harder to throw against.(categorica)\
**pff_manZone**: Specific indicator of man or zone coverage.
(categorical)\
**unblockedPressure**: Whether there was pressure from an unblocked
defender; unblocked pressure typically lowers completion probability.
(binary)\
**passTippedAtLine**: Whether the pass was tipped; this directly impacts
completion probability.(binary)

@@

```{r}
nfldata %>%
  distinct(timeToSack)
```

```{r}
nfldata_viz <- subset(
  nfldata,
  !is.na(dropbackType) & dropbackType != "UNKNOWN" & 
    passResult %in% c("C", "I")
)

ggplot(nfldata_viz, aes(x = dropbackType, fill=passResult)) + 
  geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
  fill='passResult', title = "passResult 
  in different dropbackType") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c('Complete', 
  'Incomplete'))
```

@@

```{r}
nfldata_viz <- nfldata_viz %>%
  mutate(
    dropbackType = ifelse(dropbackType %in% c("Traditional", "Rollout"), dropbackType, "Other")
  )

ggpairs(
  nfldata_viz,
  columns = c("dropbackType", "dropbackDistance" ),
  aes(color = as.factor(passResult)), # Treat passResult as a factor
  upper = list(continuous = "points"),
  lower = list(continuous = "points")
) +
  labs(
    title = "Pairwise Plots of Predictors Colored by Pass Result",
    color = "Pass Result"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

@@

```{r}

ggplot(nfldata, aes(x = preSnapHomeTeamWinProbability)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue") +
  facet_grid(~  as.factor(quarter))
  labs(title = "Win Probability Before Play by Quarter (Home Team)")

>>>>

**Outcome Variable**: `passResult` (Binary: Complete/Incomplete),
indicating whether the pass was successfully completed.

**Primary Predictors**:

-   `dropbackDistance` (Continuous).The distance the QB dropped back
    (yards) behind the center after the snap (numeric)

-   `dropbackType` (Categorical).The type of drop back after the snap by
    the QB (Traditional, Designed Rollout, Scramble, Scramble Rollout,
    Designed Rollout Left, Designed Rollout Right, Scramble Rollout
    Left, Scramble Rollout Right, Designed Run, QB Draw, Rollout, text)
    (categorical)

**Contextual Variables:**

-   `timeToThrow` (Continuous). The time (secs) elapsed between snap and
    pass (numeric)

-   `passLength` (Continuous). The distance beyond the LOS that the ball
    traveled not including yards into the endzone. If thrown behind LOS,
    the value is negative. (numeric)

**Potential Interaction Terms:**

-   `dropbackDistance* dropbackType` : Different types of dropbacks
    (e.g., "Rollout" vs. "Traditional") may influence the effectiveness
    of a specific dropback distance.

@@

```{r}
# Clean and transform dataset
nfldata_1 <- nfldata %>%
  # Select relevant variables
  select(dropbackDistance, dropbackType, timeToThrow, passLength, passResult) %>%
  # Filter for valid pass results
  filter(
    passResult %in% c("C", "I"),
    !is.na(dropbackType) & dropbackType != "UNKNOWN",
    !is.na(dropbackDistance),
    !is.na(timeToThrow),
    !is.na(passLength)
  ) %>%
  # Convert variables to factors
  mutate(
    dropbackType = as.factor(dropbackType),
    passResult = as.factor(passResult)
  )
glimpse(nfldata_1)

```{r}
ggplot(nfldata_1, aes(x = dropbackType, fill=passResult)) + 
  geom_bar() + labs(x="dropbackType", y="dropbackDistance", 
  fill='passResult', title = "passResult
  in dropbackType") + theme(axis.text.x = 
  element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_fill_discrete(labels=c("Complete", "Incomplete"))
```

@@

```{r}
table(nfldata_1$dropbackType)
```

```{r}
nfldata_1_viz <-nfldata_1 %>%
  mutate(
    dropbackCategory = case_when(
      dropbackType == "TRADITIONAL" ~ "Traditional",
      dropbackType %in% c("DESIGNED_ROLLOUT_LEFT", "DESIGNED_ROLLOUT_RIGHT") ~ "Designed Rollout",
      dropbackType %in% c("SCRAMBLE", "SCRAMBLE_ROLLOUT_LEFT", "SCRAMBLE_ROLLOUT_RIGHT") ~ "Scramble",
      TRUE ~ NA_character_  # Handle unexpected values
    )
  )

ggpairs(
  nfldata_1_viz,
  columns = c("dropbackCategory", "dropbackDistance" ),
  aes(color = as.factor(passResult)), # Treat passResult as a factor
  upper = list(continuous = "points"),
  lower = list(continuous = "points")
) +
  labs(
    title = "Pairwise Plots of Predictors Colored by Pass Result",
    color = "Pass Result"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, hjust = 0.5, face = "bold"),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

@@

```{r}
ggplot(nfldata, aes(x = preSnapHomeTeamWinProbability)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue") +
  facet_grid(~  as.factor(quarter))
  labs(title = "Win Probability Before Play by Quarter (Home Team)")
